`flippant` -- An `R` package for the automated analysis of dithionite scramblase

### 2015-09-24 inst/manuscript/manuscript.rmd
A spell checking pass;

Spectral input data is read by `flippant` using the file paths provided in the
spreadsheet. Raw spectra produced by QuantaMaster fluorometer instruments 
(Photon Technology International, Inc., Edison, New Jersey) running software 
versions FelixGX v4.1 and Felix32 v1.20, respectively, as well as a simple 
generic tab-delimited spectrum representing file format with time (in seconds) 
and fluorescence intensity representing columns are currently supported. 
File type determination is handled algorythmically.

@@

`flippant`s capabilities are demonstrated here by reanalyzing data underlying 
Figure 4C/D in @ploier_perturbed_submitted. The following work flow in a first 
step plots the corresponding spectral traces (Figure 1A). Note that alorithmic
alignment to the time of dithionite addition is employed (default) as well as 
custom trimming of the time axis to account for divergent acquisition 
times between experiments.
```{r, eval = FALSE}
# Plot spectral traces (Figure 1A)
scramblaseAssayTraces("inputTable.txt", timeMax = 350)
```
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "**`flippant`-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** (A) represents spectral traces as produced by a call to the function `scramblaseAssayTraces`, while (B) shows protein to phospholipid ration (PPR) plots for the data using a x-axis trimmed to a maximum value of 2 (see text for details)."}
setwd(tmpDir)
figure1 <- addpanel(
  figure = figure1,
  topPanel = 1,
  bottomPanel = 3,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayTraces("inputTable.txt", timeMax = 350) + theme_bw())
figure1 <- addpanel(
  figure = figure1,
  topPanel = 4,
  bottomPanel = 5,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayPlot("inputTable.txt") + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])))
grid.draw(figure1)
```

@@

Mean $\tau$ values (or fit constants $\alpha$) correspond with `r ppr_results_tau_average_print` for `r ppr_results_tau_average_names`, 
respectively, very well to the prior manual analyis by 
@ploier_perturbed_submitted, where $\tau \approx 0.12$ was reported for mustant 
and $\tau \approx 0.26$ for wild type rhodopsin. Based on the manual analysis of
the data we conclude in @ploier_perturbed_submitted that the mutant rhodopsins
investigated are reconsituted into proteoliposomes in monomeric form, contrary 
to the wild type, which is reconstituting as a homodimer. `flippant` analysis of
the data, shows identical trends and close matches of the results, supporting 
identical conclusions.

@@

analysis_timing_iqr <- round(
  diff(
    unlist(
      summary(analysis_timing)[c("lq", "uq")],
      use.names = FALSE)),
  2)
```
With `flippant` we present an integrated solution for data analysis in the 
context of dithionite scramblase assays. Requiring only basic familiarity with 
`R` [@r_core_team_r:_2015] as an environment for statistical computing and
graphics, scientists can quickly analyse such data and arrive at 
publication-grade graphics that offer extensive facilities for individual 
optimization and adaptation [@wickham_ggplot2_2009]. On a laptop system with 
`r parallel::detectCores()` cores and 8 GB RAM (no SSD), running `Linux` and 
`R` version `r paste(R.version[c("major","minor")], collapse = ".")`, the
statements generating equivalents to the plots and calculations presented here
take a mere `r analysis_timing_median` seconds (median; 
$\text{IQR}=`r analysis_timing_iqr`~\text{s}$; $n=`r analysis_timing_n`$).

Results from `flippant`-driven analysis track well prior manual analysis. 
Deviations observed may stem from algorithmic differences such as the fitting 
routine used (`flippant` uses the Nash variant of the Marquardt 
nonlinear least squares algorithm as implemented by @nash_nlmrt:_2014) and 
general handling of significant digits and rounding.

>>>>

Spectral input data is read by `flippant` using the file paths provided in the
spreadsheet. Raw spectra produced by QuantaMaster fluorometer instruments 
(Photon Technology International, Inc., Edison, New Jersey) running software 
versions FelixGX v4.1 and Felix32 v1.20, respectively, as well as a simple 
generic tab-delimited spectrum representing file format with time (in seconds) 
and fluorescence intensity representing columns are currently supported. 
File type determination is handled algorithmically.

@@

`flippant`s capabilities are demonstrated here by reanalyzing data underlying 
Figure 4C/D in @ploier_perturbed_submitted. The following work flow in a first 
step plots the corresponding spectral traces (Figure 1A). Note that algorithmic
alignment to the time of dithionite addition is employed (default) as well as 
custom trimming of the time axis to account for divergent acquisition 
times between experiments.
```{r, eval = FALSE}
# Plot spectral traces (Figure 1A)
scramblaseAssayTraces("inputTable.txt", timeMax = 350)
```
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "**`flippant`-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** (A) represents spectral traces as produced by a call to the function `scramblaseAssayTraces`, while (B) shows protein to phospholipid ration (PPR) plots for the data using a x-axis trimmed to a maximum value of 2 (see text for details)."}
setwd(tmpDir)
figure1 <- addpanel(
  figure = figure1,
  topPanel = 1,
  bottomPanel = 3,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayTraces("inputTable.txt", timeMax = 350) + theme_bw())
figure1 <- addpanel(
  figure = figure1,
  topPanel = 4,
  bottomPanel = 5,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayPlot("inputTable.txt") + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])))
grid.draw(figure1)
```

@@

Mean $\tau$ values (or fit constants $\alpha$) correspond with `r ppr_results_tau_average_print` for `r ppr_results_tau_average_names`, 
respectively, very well to the prior manual analysis by 
@ploier_perturbed_submitted, where $\tau \approx 0.12$ was reported for mutant 
and $\tau \approx 0.26$ for wild type rhodopsin. Based on the manual analysis of
the data we conclude in @ploier_perturbed_submitted that the mutant rhodopsins
investigated are reconstituted into proteoliposomes in monomeric form, contrary 
to the wild type, which is reconstituting as a homodimer. `flippant` analysis of
the data, shows identical trends and close matches of the results, supporting 
identical conclusions.

@@

analysis_timing_iqr <- round(
  diff(
    unlist(
      summary(analysis_timing)[c("lq", "uq")],
      use.names = FALSE)),
  2)
```
With `flippant` we present an integrated solution for data analysis in the 
context of dithionite scramblase assays. Requiring only basic familiarity with 
`R` as an environment for statistical computing and graphics 
[@r_core_team_r:_2015], scientists can quickly analyse such data and arrive at 
publication-grade graphics that offer extensive facilities for individual 
optimization and adaptation [@wickham_ggplot2_2009]. On a laptop system with 
`r parallel::detectCores()` cores and 8 GB RAM (no SSD), running `Linux` and 
`R` version `r paste(R.version[c("major","minor")], collapse = ".")`, the
statements generating equivalents to the plots and calculations presented here
take a mere `r analysis_timing_median` seconds (median; 
$\text{IQR}=`r analysis_timing_iqr`~\text{s}$; $n=`r analysis_timing_n`$).

Results from `flippant`-driven analysis track prior manual analysis well. 
Deviations observed may stem from algorithmic differences such as the fitting 
routine used (`flippant` uses the Nash variant of the Marquardt 
nonlinear least squares algorithm as implemented by @nash_nlmrt:_2014) and 
general handling of significant digits and rounding.
